---
title: "INLA Hurdle Gamma Example"
author: "Simon Brewer"
date: "1/30/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r message = FALSE, warning=FALSE}
library(INLA)
library(ggplot2)
library(tidyverse)
```

## Make up some data

- Generate `y`: binary outcomes with fixed probability ($p=0.7$)
- Convert to Gamma distributed variable by adding appropriate noise with fixed scale ($\theta = 0.4$)
- Convert to `z`: binary outcome for modeling
```{r}
set.seed(1)
x <- 1:150
y <- rbinom(length(x), size = 1, prob = 0.7)
y <- y * rgamma(length(x), shape = 0.4)
z <- ifelse(y > 0, 1, 0)
d <- data.frame(x, y, z)

table(pop_tox_2$rsei.score == 0)
## fix so 1 is 0
pop_tox_2$rsei.score <- ifelse(pop_tox_2$rsei.score == 1, 0, pop_tox_2$rsei.score - 1)
pop_tox_2$rsei_bin <- ifelse(pop_tox_2$rsei.score > 0, 1, 0)

pop_tox_2
```

```{r}
knitr::kable(head(d))
```

```{r}
p <- ggplot(d, aes(x, y, colour = as.factor(z))) + geom_point()
print(p)
```

## GLMs 

Make two independent GLMS: binomial (using `z`) and Gamma (using `y`)

```{r}
m1 <- glm(rsei_bin ~ 1, data = pop_tox_2, family = binomial(link = logit))
m2 <- glm(rsei.score ~ 1, data = subset(pop_tox_2, rsei_bin == 1),
          family = Gamma(link = inverse))

hist(test$rsei.score^-1)

hist(d$y)
hist(log(test$rsei.score))

test <- subset(pop_tox_2, rsei_bin >= 1)
test_2 <- subset(d, y >= 1)
test$rsei.score_log <- log(test$rsei.score)


summary(test_2$y)
summary(test$rsei.score)
m3 <- glm(z ~ 1, data = d, family = binomial(link = logit))
m4 <- glm(y ~ 1, data = subset(d, z == 1), family = Gamma(link = log))

#check model
plot(m2, cex = 0.4)
plot(m4, cex = 0.4)

```

To check if they have recovered the original parameters, we need to back transform the intercepts. First, back-transform the binomial using `plogis()`. Should give a result close to $p = 0.7$:

```{r}
(bin_coef <- plogis(coef(m1)[[1]]))
```

Now back transform the Gamma intercept. This is the estimate of scale (should be close to 0.4), and needs `exp()`:

```{r}
(gamma_coef <- exp(coef(m2)[[1]]))
```

We can also get confidence intervals on both of these using `confint`:

```{r}
plogis(confint(m1))
exp(confint(m2))

hist()

```

## INLA

Now let's run the same models as a joint model in INLA. The data setup requires making a set of duplicate outcomes, the first set will have the binary outcome (`z`) and the second the Gamma (`z`). They will also have matching covariates (here only the intercept estimates). We create a duplicate for each covariates, but using `NA` values for the other outcome. This will probably be clearer when we do it.

- Get the number of obs (we'll need this a lot)

```{r}
n = nrow(d)

n = nrow(pop_tox_2)
```

- Next, generate two vectors for the outcomes. Note for the Gamma outcome, all 0's are set to `NA`s:

```{r}
z = as.vector(pop_tox_2$rsei_bin > 0)
y = ifelse(z == 1, pop_tox_2$rsei.score, NA)

#z <- as.vector(d$y > 0)
#y <- ifelse(z == 1, d$y, NA)
```

- Now generate repeats for the outcomes, by concatenating the values with NA's. As the first model will be the binomial, we concatenate `z`, then `NA`. For the second Gamma model, we concatenate `NA` then `y`:

```{r}
nothing1 <- rep(NA, n)
nothing2 <- rep(NA, n)

zNA = as.vector(c(z, nothing1))
yNA = as.vector(c(nothing2, y))
```

- Now stick these together in a two-column array:

```{r}
outcome.matrix <- matrix(c(z, y), ncol = 2)
outcome.matrix <- matrix(c(zNA, yNA), ncol=2)

print(outcome.matrix)
```

- Next generate two vectors to capture the intercepts. Again, we concatenate with `NA` values (and use `1`s for the intercept):

```{r}
# mu = as.factor(c(rep(1,length(z)), rep(2,length(y))))
mu_z <- c(rep(1, n), nothing1) # Binomial 
mu_y <- c(nothing2, rep(1, n)) # Gamma 
```

- Any covariates have to be constructed in the same way. If we have two (`x1` and `x2`):

```{r eval=TRUE}
x1 <- rnorm(n)
x2 <- rnorm(n)

x1_z <- c(x1, nothing1) # Binomial 
x1_y <- c(nothing2, x1) # Gamma 

x2_z <- c(x2, nothing1) # Binomial 
x2_y <- c(nothing2, x2) # Gamma 
```

- Now compile the whole thing into a list

```{r}
data_hg <- list(outcome.matrix = outcome.matrix, 
                mu_z = mu_z, mu_y = mu_y,
                x1_z = x1_z, x1_y = x1_y,
                x2_z = x2_z, x2_y = x2_y)
```

## Model

```{r}
f_hg <- outcome.matrix ~ mu_z + mu_y - 1 

res1 <- inla(f_hg, family = c("binomial", "gamma"), data = data_hg,
             #control.predictor=list(link=...),
             verbose=FALSE) 

summary(res1)
```


Now check the INLA estimates of the two parameters. First the binomial ($p=0.7$):

```{r}
plogis(res1$summary.fixed$mean[1])
```
```{r}
res1$marginals.fixed$mu_z %>%
  as.data.frame() %>%
  mutate(x = plogis(x)) %>%
  ggplot(aes(x = x, y = y)) +
  geom_line()
```

And the Gamma ($\theta=0.4$):

```{r}
exp(res1$summary.fixed$mean[2])
```

```{r}
res1$marginals.fixed$mu_y %>%
  as.data.frame() %>%
  mutate(x = exp(x)) %>%
  ggplot(aes(x = x, y = y)) +
  geom_line()
```
